---
title: "Expected IL for Uniswap v3 positions"
author:
- 0xAlbert.eth
- folch.eth
- 0xOli
date: \today
abstract: ""
pdf_document: null
geometry: "margin=45mm"
output: 
  pdf_document:
      toc: true
      number_sections: true
      toc_depth: 2
      keep_md: false
      keep_tex: false
      template: NULL
      #pandoc_args: [
      #  --template=extras/haptic_template.tex
      #]
html_document: default
tables: yes
latex_engine: pdflatex
header-includes:
    - \usepackage{float}
    - \usepackage{pdfpages}
    - \usepackage{tabu}
    - \usepackage{lipsum}
    - \usepackage{booktabs}
    - \usepackage[justification=raggedright,labelfont=bf,singlelinecheck=false]{caption}
    - \usepackage{array}
    - \usepackage{xcolor} 
    - \usepackage{color, colortbl}
    - \usepackage{amsmath}
    - \usepackage{mathtools,mathptmx}
    - \usepackage{tabularx}
    - \usepackage{background}
    - \usepackage[english]{babel}
    - \usepackage{csquotes}                
    - \usepackage[style=alphabetic, backend=bibtex]{biblatex}
    - \bibliography{bibliography/haptic.bib}
    - \usepackage{tikz}
    - \usepackage[font=large,labelfont=bf]{caption}
    - \usetikzlibrary{shapes,positioning}
    - \usepackage{wrapfig}
    - \usepackage{eso-pic,graphicx,transparent}
    - \DeclareUnicodeCharacter{2212}{-}
    - \backgroundsetup{pages={some},contents={}, opacity={0.3}, color={gray}}
    - \usepackage{multirow}
    - \usepackage{caption}
    - \newcommand{\Tau}{\textstyle{\mathcal{T}}}
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "pdf") })
---
\captionsetup{font=small}
\captionsetup[table]{labelformat=empty}
\captionsetup[table]{labelfont=bf}
```{r results="asis", warning=FALSE, echo=FALSE}
options(knitr.kable.NA = '') 
options(scipen=999) # avoid scientific notation
options(width = 60)

path = getwd()
source(paste(path, "/config.r", sep=""))
source(paste(path, "/uil.r", sep=""))

source(paste(path, "/eil.r", sep=""))
source(paste(path, "/static.r", sep=""))

T = 168
P = S0

chunk_one <- run_calc(mu, sigma, T) 
chunk_two <- run_calc(mu, sigma,  T * 2) 
df_uil_one <- run_uil_calc(ranges, sigma, T)

```
\newpage

\section{Introduction}

Impermament loss (IL) or divergence-loss, is a phenomena experienced by LP providers on almost all types of AMM platforms, especially when providing liquidity for non-mean reverting assets. The effects of IL have been studied extensively because of their negative impact on LP providers' returns \cite{LOE} \cite{DZ}. When modelling an asset's price with a geometric brownian motion (GBM), IL can be estimated numerically by using iterative techniques such as the Monte Carlo method. The result of this estimation is referred to as "expected impermanent loss" (EIL). Previous work wrongly concluded that the computation of EIL for Uniswap v3 positions is analytically intractable, due to the complexity of the math expressions involving the GBM \cite{GSL}. However, a recent paper provided a framework for the analytical computation of EIL for Uniswap v3 positions \cite{CV3}. We implement the ideas presented in the paper and compare the results with the results obtained from the Monte Carlo simulation.

\subsection{Simulation parameters}
\label{simpar}

  - Number of simulations: `r n_sim` 
  - Total time: `r T` (Hours)
  - Initial price: `r S0` (Units of numéraire asset)
  - Drift: `r mu`
  - Standard deviation: `r sigma`

\newpage

\section{Technical details}

\subsection{Expected IL}
\label{eilint}

The integrals in the analytical computation of EIL are calculated using the "_integrate_" R function, which is based on QUADPACK routines dqags and dqagi by R. Piessens and E. deDoncker–Kapenga \cite{PIE}.

```{r a, eval=FALSE}

function(f, lb, ub, P, Pa, Pb, mu, sigma, T) {

    integral <- integrate(
        f,
        # Lower integration bound
        lb, 
        # Upper integration bound
        ub, 
        subdivisions = 5000,
        abs.tol = 1e-5, 
        rel.tol = 1e-7,  
        stop.on.error = F,
        T, 
        P, 
        Pa,    
        Pb,
        mu,
        sigma
    )$value;

    coefficient <- (1 / (sigma * sqrt(2 * pi * T)))
    return(coefficient * integral)
}

```

\newpage

\subsection{Time in-the-money}
\label{itimint}

 The two-dimensional integrals for the computation of "_time in-the-money_" ($\Tau$) are calculated using the dblquad function from the pracma library. 

```{r b, eval=FALSE}

function(P, Pa, Pb, mu, sigma, T) {

    lower_bound <- log(Pa / P)
    upper_bound <- log(Pb / P)

    f <- function(x, y, mu, sigma) {
        t <- y
        first_term <- -(x - (mu - (sigma^2 / 2)) * t)^2
        second_term <- (2 * t * (sigma^2))
        exp(first_term / second_term) / sqrt(t)
    }

    integral <- dblquad(
        f,
        # Lower integration bound 
        log(Pa / P), 
        # Upper integration bound
        log(Pb / P), 
        0, 
        T,
        tol=1e-5,
        subdivs=5000, 
        mu=mu, 
        sigma=sigma
    )

    result <- (1 / (sigma * (sqrt(2 * pi)))) * integral[1]
    
    return(result)
}


```

\newpage

\section{Simulation results} 
\label{simres}

\subsection{168 hours (7 days)}

```{r, echo=FALSE, warning=FALSE, results='asis'}

collapse_rows_dt <- data.frame(
    C1 = chunk_one[1:nrow(chunk_one), 1],
    C2 = chunk_one[1:nrow(chunk_one), 2],
    C3 = chunk_one[1:nrow(chunk_one), 3],
    C4 = chunk_one[1:nrow(chunk_one), 4],
    C5 = chunk_one[1:nrow(chunk_one), 5],
    C6 = chunk_one[1:nrow(chunk_one), 6],
    C7 = chunk_one[1:nrow(chunk_one), 7],
    C8 = chunk_one[1:nrow(chunk_one), 8],
    C9 = chunk_one[1:nrow(chunk_one), 9]
)

sigmaMsg <- paste("$\\\\sigma = $ ", sigma, sep="")
sigmaMsg2 <- paste("$\\\\sigma = $ ", sigma * 2, sep="")

collapse_rows_dt %>%
kbl(booktabs = T, 
    align = "c", 
    longtable = T, 
    escape = FALSE, 
    col.names = c( "$\\mu$", "r", "CE", "$\\Tau$", "EIL", "IL","$\\Tau$", "EIL", "IL"),
    caption = "Table 1: Results for setting $\\mu$ = 7.5e−5, 1.5e−4 and $\\sigma$ = 0.001 and $\\sigma$ = 0.002 for 7 days (T = 168 hours) of providing liquidity with initial price \\$1000."
    ) %>%  kable_styling(font_size = 9) %>%
    column_spec(1, bold=F) %>%
        collapse_rows(columns = 1, latex_hline = "major", valign = "middle")  %>%
            add_header_above(c(" ", " ", " ", "Formula" = 2, "Simulation" = 1, "Formula" = 2, "Simulation" = 1), escape = FALSE)  %>%
                add_header_above(escape =FALSE, c(" ", " ", setNames(3,sigmaMsg), setNames(4,sigmaMsg2)))   %>%
                    kable_styling(font_size = 9, latex_options = c("hold_position", "repeat_header")) %>%
                        print()

```

\subsection{7200 hours (30 days)}

```{r, echo=FALSE, warning=FALSE, results='asis'}
collapse_rows_dt <- data.frame(
    C1 = chunk_two[1:nrow(chunk_two), 1],
    C2 = chunk_two[1:nrow(chunk_two), 2],
    C3 = chunk_two[1:nrow(chunk_two), 3],
    C4 = chunk_two[1:nrow(chunk_two), 4],
    C5 = chunk_two[1:nrow(chunk_two), 5],
    C6 = chunk_two[1:nrow(chunk_two), 6],
    C7 = chunk_two[1:nrow(chunk_two), 7],
    C8 = chunk_two[1:nrow(chunk_two), 8],
    C9 = chunk_two[1:nrow(chunk_two), 9]
)

sigmaMsg <- paste("$\\\\sigma = $ ", sigma, sep="")
sigmaMsg2 <- paste("$\\\\sigma = $ ", sigma * 2, sep="")

collapse_rows_dt %>%
kbl(booktabs = T, 
    align = "c", 
    longtable = T,
    escape = FALSE, 
    col.names = c( "$\\mu$", "r", "CE", "$\\Tau$", "EIL", "IL","$\\Tau$", "EIL", "IL"),
    caption = "Table 2: Results for setting $\\mu$ = 7.5e−5, 1.5e−4 and $\\sigma$ = 0.001 and $\\sigma$ = 0.002 for 30 days (T = 720 hours) of providing liquidity with initial price \\$1000."
    ) %>%  kable_styling(font_size = 9) %>%
    column_spec(1, bold=F) %>%
        collapse_rows(columns = 1, latex_hline = "major", valign = "middle")  %>%
            add_header_above(c(" ", " ", " ", "Formula" = 2, "Simulation" = 1, "Formula" = 2, "Simulation" = 1), escape = FALSE)  %>%
                add_header_above(escape =FALSE, c(" ", " ", setNames(3,sigmaMsg), setNames(4,sigmaMsg2)))   %>%
                    kable_styling(font_size = 9, latex_options = c("hold_position", "repeat_header")) %>%
                        print()

static_expected_prices <- get_expected_prices(P, mu)
data <- static_expected_prices$data

df_expected_prices <- data.frame(data)
colnames(df_expected_prices) <- c("$\\mu$", "$\\Tau$ = 168", "$\\Tau$ = 720")
df_expected_prices <-  setNames(df_expected_prices, c("$\\mu$", "$\\Tau$ = 168", "$\\Tau$ = 720"))
row.names(df_expected_prices) <- NULL

df_expected_prices %>%
kbl( escape = F, caption = paste("Table 3: Expected price of the asset, for 7 ($\\Tau$ = 168 hours) and 30 days ($\\Tau$ = 720 hours)."), booktabs = T) %>%
    kable_styling(latex_options = "HOLD_position", position="left") %>%
    kable_classic(full_width = F, html_font = "Cambria") %>%
      print()

df_uil_one %>%
    kbl(booktabs = T, 
        align = "l", 
        longtable = T,
        escape = FALSE, 
        caption = "Table 3 - Unit impermanent loss per liquidity (UIL)"
    ) %>%  
    kable_styling(font_size = 9)  %>%
       collapse_rows(columns = c(1), latex_hline = "major", valign = "middle")  %>%
            add_header_above(escape = FALSE, c(" ", "Formula " = 2, "Static replication " = 2))  %>%
                print()
```
\section{Conclusion}
\label{conclusion}

\lipsum[3-10]

\newpage

\printbibliography
